---
title: "HelicopterReport"
author: "Tristan Zanowic"
date: "4/6/2023"
output:
  pdf_document:
    toc: yes
    toc_depth: '2'
---

```{r Front Matter, include=FALSE}
# clean up & set default chunk options
rm(list = ls())
knitr::opts_chunk$set(echo = FALSE)
# packages
library(tidyverse) # for example
library(mosaic)    # for example
# user-defined functions
# inputs
data("mtcars")  # for example
```


# Project Description

*Should be in paragraph from (not bullets).  Written for a 3rd party such as your client’s boss, or an external review board.  As part of your description, consider the following:*

- *Type of study [Designed experiment, Observational study, Sample survey] and brief description of design*
- *Stage of the experiment [Planning stage or Analysis stage]*
- *Objective of study and role of this analysis*
- *Scope of project*
- *Intended use for results*

## Research Questions

*What are the overarching research questions that the client is targeting?*

**Question 1:**  

**Question 2:**  

...

*(order them by priority; RQ1 is highest priority)*


## Variables

*What is (are) possible explanatory and response variables?  Relevant notes about how each is measured/recorded.*

*A table is recommended here. *


*If design is complex, here is a good place to write a detailed description.*

# Screening Experiment


# Response Surface Experiment



# Recommendations 

*Succinct response to each question laid out in 1.2.  This a much shorter version of section 3, and focuses on conclusions rather than the analyses.*

**Question 1:**  

**Question 2:**  

*(all research questions in 1.1 and/or statistical question in 1.2 should be addressed, in the same order)*

## Considerations


# Technical Appendix  

*Detailed information and a copy of code and or software results.  Additional graphs and supporting figures may also be placed in the appendix.*


### R Script
```{r ref.label=c('Front Matter', 'Data Cleaning', 'MPG summary', 'EDA-scatterplot'), echo=TRUE, eval=FALSE}
# Reprinted code chunks used previously for analysis
```