---
title: "HelicopterReport"
author: "Tristan Zanowic, Aureo Zanon"
date: "4/6/2023"
output:
  pdf_document:
    toc: yes
    toc_depth: '2'
---

```{r Front Matter, include=FALSE}
# clean up & set default chunk options
rm(list = ls())
knitr::opts_chunk$set(echo = FALSE)
# packages
library(tidyverse) 
library(mosaic)  
library(FrF2) # tools for design and analysis of fractional factorials
library(pid)
library(MASS)
library(stats)
library(rsm)
library(kableExtra)
```


# Project Description

Paper helicopters are a fun, creative way for child care programs to get children engaged with curiosity and building. Built using a single sheet of paper and other common supplies, these helicopters are designed in a near infinite number of ways in an attempt to gain the most air time. 

To determine the optimal paper helicopter specifications in lengthening flight time, a fractional factorial design is conducted along with a response surface experiment in this research. The goal is to analyze the effects of several variables within the helicopter’s design to discover what factors lead to the longest flight time, while maintaining a budget of 60 helicopters. With this study, the results can hopefully help children build the best helicopter to stay in the air, keeping them intrigued and curious throughout their learning. 

## Research Questions


**Question 1:** What paper helicopter design factors lead to the longest flight time?

**Question 2:**  



## Variables

There are a total of six explanatory variables of interest in predicting flight time, from cutting dimensions such as wing length to add-ons like paper clips. Each variable has two levels to be analyzed (listed in the description column of the table below).

```{r}
#variable table
vartable = data.frame(Variable = c("Wing length","Body length","Body width","Paper clips","Folded wings","Folded body","Time"), Type = c("Explanatory","Explanatory","Explanatory","Explanatory","Explanatory","Explanatory", "Response"), Description = c("Length of wings (70mm or 120mm)", "Length of body (70mm or 120mm)", "Width of body (35mm or 50mm)", "No. of paper clips (0 or 2)", "Are wings folded (no or yes)", "Is the body folded (no or yes)", "Time in air (seconds)"))
   

vartable %>%
  kbl(caption = "Variable Table") %>%
  kable_classic(full_width = F, html_font = "Cambria") 
```


# Screening Experiment

To begin analyzing the six variables of interest, a fractional factorial design is used to screen our predictors. To identify the key variables at play in lengthening paper helicopter flight time, a resolution VI factorial design is used, equating to 32 screening runs, with four additional runs for center points. Within these runs, helicopters were dropped from approximately 15 feet with variations in design specifications across each trial, and the time in air was recorded. One observation that was noted during these trials was the common appearance of tumbling, which seemed to occur more often when paper clips were used. Regardless, no obvious trends related to flight time were observed during the tests.

From these results of these 32 runs, a model can be fitted including our six variables of interest, along with every 2-way interaction. To reduce this complex model, step-wise backwards selection was used to drop the least significant terms, and a summary of this model can be seen in the table below. As seen in this table, the most significant terms appear to be folded body and paper clips, with both having a negative effect on flight time. Many interaction terms (that include folded body and paper clips) being significant as well. Other variables of interest, such as body length, body width, and folded wings, appear to have negligible effects on their own (p-values > .6).

```{r}
HeliDesign <- read.csv("fracfact.csv")
#store then omit center points 
CP <- HeliDesign[33:36,]
HeliDesign <- HeliDesign[1:32,]


# main effects and all 2-way interactions
HeliMod_2way <- lm(time ~(.)^2, data = HeliDesign)

# looking for contributing effects
#pid::paretoPlot(HeliMod_2way)
# reduced model after stepwise backwards selection (alpha = 0.1)**
step <- stepAIC(HeliMod_2way, direction = "both", trace = FALSE)

HeliMod_reduced <- step
sumHeli <- summary(HeliMod_reduced)
screeningtable <- as.data.frame(round(sumHeli$coefficients,3))
names(screeningtable) <- c("Coefficient","Standard Error", "T-Test Stat", "P-value")

screeningtable %>%
  kbl(caption = "Reduced Fractional Factorial Model Results") %>%
  kable_classic(full_width = F, html_font = "Cambria") 
```

The pareto plot below further emphasizes these effects, showcasing which variables (including interactions) have the greatest effect (both negative and positive) on flight time. In agreement with the model results above, having a folded body had the largest effect on flight time, as the presence of a folded body negatively effected time in the air. Following this, the interaction between body length and paper clips had the largest positive effect, as the inclusion of a longer body (120mm) and two paper clips generally led to longer flight time. However, paper clips on its own had a large negative impact on flight time, the third largest effect.

```{r}
paretoPlot(HeliMod_reduced)
```


```{r}
# effect plots
MEPlot(HeliMod_reduced) # main effects
```

```{r}
IAPlot(HeliMod_reduced) # interactions
cubePlot(HeliMod_reduced, eff1 = "winglength", eff2 = "paperclips", 
         eff3 = "foldedbody",
         main = "Cube Plot for Distance to Optimal Helicopter")
```

# Response Surface Experiment

Upon analyzing the results of the screening process, we chose to further optimize the three most significant (not including interactions) explanatory variables: **folded body, paper clips, and wing length**. To do so, a box-behnken design is used for our response surface experiment, equating to 16 optimization runs. 

Following measuring the time of these 16 helicopters, we can fit a model to describe the effects of the three variables of interest. Most significantly (p-value < .001), paper clips appeared to have the largest effect on flight time, as the inclusion of these paper clips leads to about a one second drop in paper helicopter flight time, after accounting for folded body and wing length.

```{r}
##boxbehnken design
bb_design_1 <- bbd(
  k  = 3,            # Number of factors,
  n0 = 4,            # Number of center points,
  block = FALSE,     # Consider blocks or not in the design 
  randomize = FALSE)

names <-c("foldedbody","paperclips","winglength")

designbb <- bb_design_1[,3:5]
#setNames(designbb,names)
#fwrite(designbb,"~/Desktop/PSU/SeniorYear/Spring Semester/STAT470W/HeliFFD/BBDesign.csv")

#response surface design
HeliD <- read.csv("BBDesign.csv")
model<-rsm(time~ FO(foldedbody,paperclips,winglength), data=HeliD) #***


sumBBHeli <- summary(model)
BBtable <- as.data.frame(round(sumBBHeli$coefficients,3))
names(BBtable) <- c("Coefficient","Standard Error", "T-Test Stat", "P-value")

BBtable %>%
  kbl(caption = "Box Behnken Optimization Model Results") %>%
  kable_classic(full_width = F, html_font = "Cambria") 
```


Following analysis of the model we fit for our response surface experiment, ____ reveal how each of our three variables interacts with each other to maximize flight time. To begin, in the ___ with paper clips and folded body our flight time is maximized when our helicopter body is not folded, and paper clips are not included. From the other two ___, the results from our model are upheld, showing that the optimal helicopter design has longer wings, no paper clips, and no folded body. 

```{r}
# countour plots
par(mfrow=c(1,3))
#contour(model,~foldedbody+paperclips+winglength, 
       # image=TRUE, at=summary(model$canonical$xs))

dev.off()

par(mfrow=c(1,3))
persp(model, ~ foldedbody + paperclips, image = TRUE,
      at = c(summary(model)$canonical$xs, Block="B2"),
      theta=30,zlab="Time",col.lab=33,contour="colors")
persp(model, ~ winglength + paperclips, image = TRUE,
      at = c(summary(model)$canonical$xs, Block="B2"),
      theta=30,zlab="Time",col.lab=33,contour="colors")
persp(model, ~ winglength + foldedbody, image = TRUE,
      at = c(summary(model)$canonical$xs, Block="B2"),
      theta=30,zlab="Time",col.lab=33,contour="colors")
```

```{r}
#optimal design for longest time in the air
steepest(model)

```

# Recommendations 

*Succinct response to each question laid out in 1.2.  This a much shorter version of section 3, and focuses on conclusions rather than the analyses.*

**Question 1:**  The optimal paper helicopter design to increase flight time includes longer wings (120mm), no folded body, and no paper clips.


## Considerations

While the results of this study can likely be generalized to a classroom setting, it is important to note some limitations and additional considerations with this research. In particular, there are opportunities for human error that comes along with any building project, but especially with paper helicopters. While cutting and building these helicopters, we tried our best to consistently cut and fold in the same manner over 50 trials, but obvious limitations in human precision (and scissors) can lead to small amounts of error across helicopters. Similarly, while the helicopters were all dropped from the same spot across each trial, room factors such as the use of HVAC systems could lead to differing environmental conditions across hours or days. Regardless, with many trials, the error introduced with these limitations is ideally negligible. 

# Technical Appendix  

### R Script
```{r ref.label=c('Front Matter', 'Data Cleaning', 'MPG summary', 'EDA-scatterplot'), echo=TRUE, eval=FALSE}
# Reprinted code chunks used previously for analysis
```